---
title: "Hands-on Exercise 04"
author: "XueRong"
---

# Overview

In this hands-on exercise, we will compute and visualize spatial weights in R using geospatial data. We will focus on contiguity-based (Queen and Rook) and distance-based spatial weights matrices.

## Prerequisites

We will use the following R packages:

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr)
```

## 1. Load the Data

### Import Shapefile into R

\`\`\`{}}

``` {r}
hunan <- st_read(dsn = "/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/Hands-on_Ex/Hands-on_Ex04/data/geospatial", layer = "Hunan")
```

### Import CSV Data

``` {r}
hunan2012 <- read_csv("/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/Hands-on_Ex/Hands-on_Ex04/data/aspatial/Hunan_2012.csv")
hunan <- left_join(hunan, hunan2012)
```

## 2. Compute Contiguity-based Spatial Weights

### Queen-based Contiguity

``` {r}
wm_q <- poly2nb(hunan, queen = TRUE)
summary(wm_q)
```

### Rook-based Contiguity

``` {r}
wm_r <- poly2nb(hunan, queen = FALSE)
summary(wm_r)
```

## 3. Visualizing Spatial Weights

``` {r}
# Extract centroids of polygons
coords <- st_coordinates(st_centroid(hunan))

# Plot the base map
plot(st_geometry(hunan), border = "lightgrey")

# Plot the spatial weights with the correct coordinates
plot(wm_q, coords, pch = 19, add = TRUE, col = "red")
```

## 4. Compute Distance-based Weights

``` {r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
summary(wm_d62)
```


