---
title: "In_class_Ex05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Load Required Libraries

To begin, ensure that the necessary R packages are loaded.

```{r}
pacman::p_load(sf, sfdep, tmap, tidyverse)
```

## Step 2: Load Data

The Hunan data sets are loaded from both a shapefile and a CSV file.

```{r}
# Load the shapefile for Hunan province
hunan <- st_read(dsn = "/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/In_class_Ex/In_class_Ex05/geospatial", layer = "Hunan")

# Load the attribute data for Hunan from a CSV file
hunan2012 <- read_csv("/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/In_class_Ex/In_class_Ex05/aspatial/Hunan_2012.csv")
```

## Step 3: Merge Geospatial and Attribute Data

Use a left join to combine the shapefile and attribute datasets while retaining the spatial properties.

```{r}
hunan_GDPPC <- left_join(hunan, hunan2012) %>%
  select(1:4, 7, 15)
```

## Step 4: Plot Choropleth Map

Create a choropleth map visualizing the GDP per capita distribution across counties in Hunan province.

```{r}
tmap_mode("plot")
tm_shape(hunan_GDPPC) +
  tm_fill("GDPPC", style = "quantile", palette = "Blues", title = "GDPPC") +
  tm_layout(main.title = "Distribution of GDP per capita by county, Hunan Province",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2)
```

## Step 5: Derive Queen's Contiguity Weights

Calculate spatial weights based on Queen’s contiguity for each county.

```{r}
wm_q <- hunan_GDPPC %>%
  mutate(nb = st_contiguity(geometry),
         wt = st_weights(nb, style = "W"),
         .before = 1)
```

## Step 6: Global Moran's I Calculation

Compute the Moran's I statistic to identify the presence of spatial autocorrelation.

```{r}
moranI <- global_moran(wm_q$GDPPC, wm_q$nb, wm_q$wt)
glimpse(moranI)
```

## Step 7: Local Moran's I Calculation

Calculate local Moran's I and visualize clusters and outliers in the GDPPC data.

```{r}
lisa <- wm_q %>% 
  mutate(local_moran = local_moran(GDPPC, nb, wt, nsim = 99),
         .before = 1) %>%
  unnest(local_moran)

# Plotting the LISA map
tmap_mode("plot")
tm_shape(lisa) +
  tm_fill("ii", midpoint = NA) + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8)) +
  tm_layout(main.title = "local Moran's I of GDPPC",
            main.title.size = 2)
```

## Step 8: Hot Spot and Cold Spot Area Analysis

Perform a Hot Spot and Cold Spot Area Analysis using local Gi\* statistics.

```{r}
wm_idw <- hunan_GDPPC %>%
  mutate(nb = include_self(st_contiguity(geometry)),
         wts = st_inverse_distance(nb, geometry, scale = 1, alpha = 1),
         .before = 1)

HCSA <- wm_idw %>% 
  mutate(local_Gi = local_gstar_perm(GDPPC, nb, wts, nsim = 99),
         .before = 1) %>%
  unnest(local_Gi)

# Visualizing the hot and cold spots
tmap_mode("plot")
tm_shape(HCSA) +
  tm_fill("gi_star", midpoint = NA) + 
  tm_borders(alpha = 0.5) +
  tm_view(set.zoom.limits = c(6,8))
```
