---
title: "Hands-on_Ex02"
---

Installing and Loading the R packages

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

Spatial Data Wrangling

Importing the spatial data

```{r}
childcare_sf <- st_read("/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/Hands-on_Ex/Hands-on_Ex02/geospatial/child-care-services-geojson.geojson") %>%
  st_transform(crs = 3414)
```

```{r}
sg_sf <- st_read(dsn = "/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/Hands-on_Ex/Hands-on_Ex02/geospatial/", layer="CostalOutline")
```

```{r}
mpsz_sf <- st_read(dsn = "/Users/sharon/OneDrive - Singapore Management University/ISSS626 Data/Hands-on_Ex/Hands-on_Ex02/geospatial/", 
                layer = "MP14_SUBZONE_WEB_PL")
```

Mapping the geospatial data sets

```{r}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

```{r}
tmap_mode('plot')
```

Geospatial Data wrangling

```{r}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz_sf)
sg <- as_Spatial(sg_sf)
```

```{r}
childcare
```

```{r}
mpsz
```

```{r}
sg
```

Converting the Spatial\* class into generic sp format

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r}
childcare_sp
```

```{r}
sg_sp
```

Converting the generic sp format into spatstat’s ppp format

```{r}
childcare_ppp <- as.ppp(childcare_sf)
childcare_ppp
```

```{r}
plot(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

Handling duplicated points

```{r}
any(duplicated(childcare_ppp))
```

```{r}
multiplicity(childcare_ppp)
```

```{r}
sum(multiplicity(childcare_ppp) > 1)
```

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```

```{r}
tmap_mode('plot')
```

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

```{r}
any(duplicated(childcare_ppp_jit))
```

```{r}
# Transform the CRS to SVY21 (EPSG:3414)
sg_sf_projected <- st_transform(sg_sf, 3414)

# Check if the CRS transformation was successful
st_crs(sg_sf_projected)

```

Creating owin object

```{r}
sg_owin <- as.owin(sg_sf_projected)
```

```{r}
plot(sg_owin)
```

```{r}
summary(sg_owin)
```

Combining point events object and owin object

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]
```

```{r}
summary(childcareSG_ppp)
```

First-order Spatial Point Patterns Analysis

```{r}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 
```

```{r}
plot(kde_childcareSG_bw)
```

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
```

```{r}
childcareSG_ppp.km <- rescale.ppp(childcareSG_ppp, 1000, "km")
```

```{r}
kde_childcareSG.bw <- density(childcareSG_ppp.km, sigma=bw.diggle, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG.bw)
```

Working with different automatic badwidth methods

```{r}
bw.CvL(childcareSG_ppp.km)
```

```{r}
bw.scott(childcareSG_ppp.km)
```

```{r}
bw.ppl(childcareSG_ppp.km)
```

```{r}
bw.diggle(childcareSG_ppp.km)
```

```{r}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

Working with different kernel methods

```{r}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

Fixed and Adaptive KDE

```{r}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

```{r}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

```{r}

# Load the required libraries
library(spatstat.geom)  
library(sp)         

# Step 1: Extract pixel values (z) and grid coordinates (x, y)
x_coords <- kde_childcareSG.bw$xcol   # x coordinates of the grid
y_coords <- kde_childcareSG.bw$yrow   # y coordinates of the grid
z_values <- t(kde_childcareSG.bw$v)   # Pixel values (transposed to match coordinate orientation)

# Step 2: Create a grid of the coordinates
grid <- expand.grid(x = x_coords, y = y_coords)

# Step 3: Create a SpatialPixelsDataFrame (used for gridded data)
sp_grid <- SpatialPixelsDataFrame(points = grid, 
                                  data = data.frame(value = as.vector(z_values)),
                                  proj4string = CRS("+proj=tmerc +lat_0=1.36666666666667 +lon_0=103.833333333333 +k=1 +x_0=28001.642 +y_0=38744.572 +datum=WGS84 +units=m +no_defs"))

# Step 4: Convert to SpatialGridDataFrame
sp_grid_df <- as(sp_grid, "SpatialGridDataFrame")

# Check the structure
summary(sp_grid_df)

```

```{r}
#gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
#spplot(gridded_kde_childcareSG_bw)
```

```{R}
# Load the raster package
library(raster)

```

```{r}
kde_childcareSG_bw_raster <- raster(kde_childcareSG.bw)
```

```{r}
kde_childcareSG_bw_raster
```

```{r}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

Visualising the output in tmap

```{r}
# Load the tmap package
library(tmap)

tmap_mode("plot")  # For interactive maps

```

```{r}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("layer", palette = "viridis") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

Comparing Spatial Point Patterns using KDE

```{r}
# Load the dplyr package
library(dplyr)
```

```{r}
pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
pg <- mpsz_sf %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_sf %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_sf %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_sf %>%
  filter(PLN_AREA_N == "JURONG WEST")
```

```{r}
par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
```

```{r}
plot(tm, main = "Tampines")
```

```{r}
plot(ck, main = "Choa Chu Kang")
```

```{r}
plot(jw, main = "Jurong West")
```

```{r}
# Load the spatstat.geom package
library(spatstat.geom)

```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)
```

```{r}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

```{r}
# Load the spatstat package
library(spatstat)

```

```{r}
childcare_pg_ppp.km <- rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km <- rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km <- rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km <- rescale.ppp(childcare_jw_ppp, 1000, "km")

# Calculate bandwidths using bw.ppl()
bw_pg <- bw.ppl(childcare_pg_ppp.km)
bw_tm <- bw.ppl(childcare_tm_ppp.km)
bw_ck <- bw.ppl(childcare_ck_ppp.km)
bw_jw <- bw.ppl(childcare_jw_ppp.km)

# Perform kernel density estimation with the calculated bandwidth
kde_pg <- density(childcare_pg_ppp.km, sigma = bw_pg, edge = TRUE, kernel = "gaussian")
kde_tm <- density(childcare_tm_ppp.km, sigma = bw_tm, edge = TRUE, kernel = "gaussian")
kde_ck <- density(childcare_ck_ppp.km, sigma = bw_ck, edge = TRUE, kernel = "gaussian")
kde_jw <- density(childcare_jw_ppp.km, sigma = bw_jw, edge = TRUE, kernel = "gaussian")


# Plot the kernel density estimates
par(mfrow = c(2, 2))
plot(kde_pg, main = "Punggol")
plot(kde_tm, main = "Tampines")
plot(kde_ck, main = "Choa Chu Kang")
plot(kde_jw, main = "Jurong West")
```

```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
```

```{r}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="JUrong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

Nearest Neighbour Analysis

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

```{r}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```
